---
title: "bleach_manuscript_code"
author: "Kelly Waters"
date: "2026-03-04"
output: html_document
---

packages
```{r}
library(tidyverse)
library(qiime2R)
library(phyloseq)
library(vegan)
library(microbiome)
library(RColorBrewer)
library(ggplot2)
library(hrbrthemes)
library(ggh4x)
library(randomForest)
library(pairwiseAdonis)
library(knitr)
library(plyr)
library(ggpubr)
library(mia)
library(miaViz)
library(NetCoMi)
library(limma)
library(sechm)
library(pheatmap)
library(stats)
library(dplyr)
library(data.table)
library(MiscMetabar)
library(circlize)
library(ggVennDiagram)
library(microeco)
library(tidyr)
library(emmeans)
library(glue)
```

read in data + filter
```{r}
set.seed(2000)
setwd("C:/soil_data/summer.soil")

#all together
phy.summer <-qza_to_phyloseq("summer-table-no-mito.qza","summer-rooted-tree.qza","summer-taxonomy.qza","summer-meta2.tsv")

phy.summer.filt <- prune_samples(sample_sums(phy.summer) >= 12000, phy.summer) #removing samples that have less than 12000 reads, not rarefying any 

```

alpha diversity 
calculate 
```{r}
still_fam <- tax_glom(phy.still, taxrank = "Family")
aug_richness_all <- estimate_richness(aug_fam, measures = c("Observed", "Shannon", "InvSimpson"))
```


violin alpha diversity plots 
only stills
```{r}
#extract the data from plot_richness
richness_df <- plot_richness(phy.still, x = "Sampling.point", measures = "Observed")$data

violin.asvs.test4 <- ggplot(richness_df, aes(x = Sampling.point, y = value, fill = Treatment)) +
  geom_violin(
    position = position_dodge(width = 0.9),
    alpha = 0.6,
    trim = FALSE
  ) +
  geom_boxplot(
    width = 0.15,
    position = position_dodge(width = 0.9),
    alpha = 0.8
  ) +
  geom_jitter(
    aes(fill=Treatment),
    shape = 21,
    color = "black",
    position = position_jitterdodge(jitter.width = 0.1, dodge.width = 0.9),
    size = 1.5,
    alpha = 0.9,
    show.legend = FALSE
  ) +
  theme_classic() +
  scale_fill_brewer(palette = "Paired") +
  scale_color_brewer(palette = "Paired") +
  xlab("Sampling Point") +
  ylab("ASV Richness")
violin.asvs.test4

#family level
still_fam <- tax_glom(phy.still, taxrank = "Family")
richness_df_fam <- plot_richness(still_fam, x = "Sampling.point", measures = "Observed")$data

violin.fam.still <- ggplot(richness_df_fam, aes(x = Sampling.point, y = value, fill = Treatment)) +
  geom_violin(
    position = position_dodge(width = 0.9),
    alpha = 0.6,
    trim = FALSE
  ) +
  geom_boxplot(
    width = 0.15,
    position = position_dodge(width = 0.9),
    alpha = 0.8
  ) +
  geom_jitter(
    aes(fill=Treatment),
    shape = 21,
    color = "black",
    position = position_jitterdodge(jitter.width = 0.1, dodge.width = 0.9),
    size = 1.5,
    alpha = 0.9,
    show.legend = FALSE
  ) +
  theme_classic() +
  scale_fill_brewer(palette = "Paired") +
  scale_color_brewer(palette = "Paired") +
  xlab("Sampling Point") +
  ylab("Bacterial Family Richness")
violin.fam.still

still_richness_all2 <- estimate_richness(still_fam, measures = c("Observed", "Shannon", "InvSimpson"))
still_meta <- read.delim("still-meta.txt") #metadata file with Sample.ID and sample names with . not -
still_richness4 <- setDT(still_richness_all2, keep.rownames = "Sample.ID")
still.div.meta <- merge(still_richness4, still_meta, by='Sample.ID', all.x=TRUE)
still.div.meta2 <- merge(still_richness4, still_meta, by='Sample.ID', all.x=TRUE)
file.path <- "C:/soil_data/summer.soil/plots/still_alpha_wmeta.csv"
write.csv(still.div.meta2, file=file.path)
```

anova for alpha diversity
```{r}
#aggregate to get the average of the body site, treatment, and day
#soil_alpha_ag <- aggregate(cbind(Observed, Shannon, InvSimpson)~Treatment+Sampling.point+Sampling.location, still.div.meta2, FUN=mean)

##splitting data frame to each body site 

#stillnew <- split(still.div.meta2, still.div.meta2$Sampling.location)
#sidenew <- stillnew$S
#undernew <- stillnew$U
#undernew<- undernew %>% mutate(Sampling.point = factor(Sampling.point))
#sidenew <- sidenew%>% mutate(Sampling.point=factor(Sampling.point))

side_alphao <- lm(Observed~Treatment*Sampling.point, data = sidenew)
summary(side_alphao)
plot(side_alphao)

side_alphash <- lm(Shannon~Treatment*Sampling.point, data = sidenew)
summary(side_alphash)
plot(side_alphash)

side_alphais <- lm(InvSimpson~Treatment*Sampling.point, data = sidenew)
summary(side_alphais)
plot(side_alphais)

under_alphao <- lm(Observed~Treatment*Sampling.point, data = undernew)
summary(under_alphao)
plot(under_alphao)

under_alphas <- lm(Shannon~Treatment*Sampling.point, data = undernew)
summary(under_alphas)
plot(under_alphas)


under_alphais <- lm(InvSimpson~Treatment*Sampling.point, data = undernew)
summary(under_alphais)
plot(under_alphais)

```
pot-hoc testing
```{r}
contuo<-emmeans(under_alphao,~Sampling.point*Treatment)
contuo
contrast<-as.data.frame(contrast(contuo, method = "pairwise")) 
contrast %>% arrange(p.value)

contus<-emmeans(under_alphas,~Sampling.point*Treatment)
contus
contrasts<-as.data.frame(contrast(contus, method = "pairwise")) 
contrasts %>% arrange(p.value)

contuis<-emmeans(under_alphais,~Sampling.point*Treatment)
contuis
contrastis<-as.data.frame(contrast(contuis, method = "pairwise")) 
contrastis %>% arrange(p.value)

```


filter the one object by stillborn (bleach trials)
```{r}
phy.still <- subset_samples(phy.summer.filt, Size=="Still")
```

pcoa of only the stills 
color by bleach/control, shape is side or under
weighted unifrac
```{r}
ord.still = ordinate(phy.still, "PCoA", "unifrac", weighted=TRUE)

plot.ord.still.wu <- plot_ordination(phy.still, ord.still, color = "Treatment", shape="Sampling.location" )+
  geom_point(alpha = 0.7,size = 8) +
  xlab("PCoA1 (58.2%)") +
  ylab("PCoA2 (7%)") +
  theme_classic() +
  scale_color_brewer(palette = "Set1")+
  ggtitle("Weighted UniFrac Stills")+
  stat_ellipse(level=0.95)
plot.ord.still.wu$layers <- plot.ord.still.wu$layers[-1]
plot.ord.still.wu
```

permanova
```{r}
GPdist.still=phyloseq::distance(phy.still, "wunifrac")

still.adonis <- adonis2(GPdist.still ~ Treatment*Sampling.location, as(sample_data(phy.still), "data.frame"), by="terms")
still.adonis
```

comparing bleach vs control of the under samples 
```{r}
GPdist.still=phyloseq::distance(phy.still, "wunifrac")

still.adonis <- adonis2(GPdist.still ~ Treatment*Sampling.location, as(sample_data(phy.still), "data.frame"), by="terms")
still.adonis
```



re-do all of this at family level
pcoa
```{r}
still_fam <- tax_glom(phy.still, taxrank = "Family")
ord.still.fam = ordinate(phy.summer.filt, "PCoA", "unifrac", weighted=TRUE)

plot.ord.still.fam <- plot_ordination(still_fam, ord.still.fam, color = "Treatment", shape="Sampling.location" )+
  geom_point(alpha = 0.7,size = 8) +
  xlab("PCoA1 (58.2%)") +
  ylab("PCoA2 (7%)") +
  theme_classic() +
  scale_color_manual (values=c("skyblue3","violetred4"), labels = c("Bleach", "Control"))+
  guides(shape=guide_legend(title = "Location"))+
  scale_shape_discrete(labels = c("Control", "Under Carcass"))+
  stat_ellipse(level=0.95)
plot.ord.still.fam$layers <- plot.ord.still.fam$layers[-1]
plot.ord.still.fam
```

permanova
asv
```{r}
GPdist.still=phyloseq::distance(phy.still, "wunifrac")

still.adonis <- adonis2(GPdist.still ~ Treatment*Sampling.location, as(sample_data(phy.still), "data.frame"), by="terms")
still.adonis


still.adonis2 <- adonis2(GPdist.still.fam ~ Treatment*Sampling.location*Sampling.point, as(sample_data(phy.still), "data.frame"), by="terms")
still.adonis2
```

family level
```{r}
GPdist.still.fam=phyloseq::distance(still_fam, "wunifrac")

still.adonis.fam <- adonis2(GPdist.still ~ Treatment*Sampling.location, as(sample_data(still_fam), "data.frame"), by="terms")
still.adonis.fam


still.adonis.fam2 <- adonis2(GPdist.still.fam ~ Treatment*Sampling.location*Sampling.point, as(sample_data(still_fam), "data.frame"), by="terms")
still.adonis.fam2

#under only
GPdist.still.famu=phyloseq::distance(phy.still.fam.under, "wunifrac")
still.adonis.famu <- adonis2(GPdist.still.famu ~ Treatment*Sampling.point,as(sample_data(phy.still.fam.under), "data.frame"), by="terms")
still.adonis.famu

```

pairwise permanova
family
```{r}
#doing just under
phy.still.fam.under <- subset_samples(still_fam, Sampling.location=="U")

still.fam_dist_matrix <- phyloseq::distance(phy.still.fam.under, method="wunifrac")
vegan::adonis2(still.fam_dist_matrix ~ phyloseq::sample_data(ps_still.famu)$Sampling.point)

results.famu
write.csv(results.famu, "still-sampling.pointu-pairwise-wunifac.csv")
```


bar plot
```{r}
merge_less_than_top <- function(pobject, top=7){
  transformed <- transform_sample_counts(pobject, function(x) x/sum(x))
  otu.table <- as.data.frame(otu_table(transformed))
  otu.sort <- otu.table[order(rowMeans(otu.table), decreasing = TRUE),]
  otu.list <- row.names(otu.sort[(top+1):nrow(otu.sort),])
  merged <- merge_taxa(transformed, otu.list, 1)
  for (i in 1:dim(tax_table(merged))[1]){
    if (is.na(tax_table(merged)[i,2])){
      taxa_names(merged)[i] <- "Other"
      tax_table(merged)[i,1:7] <- "Other"}
  }
  return(merged)
}

family_order <- c("Chitinophagaceae", "Chthoniobacteraceae", "Comamonadaceae", "Flavobacteriaceae", "Nitrosomonadaceae", "Planococcaceae", "Sphingomonadaceae", "uncultured", "Vicinamibacteraceae", "Wohlfahrtiimonadaceae", "Xanthobacteraceae", "Other")
family_colors <- c("Chitinophagaceae"="#8DD3C7", "Chthoniobacteraceae"="#FFFFB3", "Comamonadaceae"="#BEBADA", "Flavobacteriaceae"="#FB8072", "Nitrosomonadaceae"="#80B1D3", "Planococcaceae"="#FDB462", "Sphingomonadaceae"="#B3DE69", "uncultured"="#FCCDE5" , "Vicinamibacteraceae"="#FFED6F", "Wohlfahrtiimonadaceae"="#BC80BD", "Xanthobacteraceae"="#CCEBC5", "Other"="#D9D9D9" )
still.fam.top10 <- merge_less_than_top(still_fam, top=9)

anew2 <- plot_bar(still.fam.top10,fill = "Family")+
  scale_fill_manual(values=family_colors)+
  ylab("Relative Abundance (%)")
anew2 <- anew2 + facet_wrap(~Treatment, scales="free_x") #gets rid of the empty space
anew2

still.fam.top12 <- merge_less_than_top(still_fam, top=11)

anew3 <- plot_bar(still.fam.top12,fill = "Family")+
  scale_fill_manual(values=family_colors, breaks = family_order)+
  ylab("Relative Abundance (%)")
anew3 <- anew3 + facet_wrap(~Treatment, scales="free_x") #gets rid of the empty space
anew3
```



random forest
this is side and under
using both bleach and control to predict sampling point
```{r}
set.seed(2000)
GPr.soil  = transform_sample_counts(phy.still, function(x) x / sum(x) ) #transform samples based on relative abundance
GPr.soil = filter_taxa(GPr.soil, function(x) mean(x) > 1e-5, TRUE)

Family.soil<-tax_glom(GPr.soil,"Family")

otu.soil<-t(otu_table(Family.soil))

df_soil<-as.data.frame(otu.soil)
taxa_names(Family.soil)
tax_table(Family.soil)
identical(colnames(df_soil), taxa_names(Family.soil))
colnames(df_soil)<-as.data.frame(tax_table(Family.soil))$Family
df_soil


predictors.soil=t(otu_table(Family.soil))
response.soil <- as.factor(sample_data(Family.soil)$Sampling.point)
rf.data.soil <- data.frame(response.soil, df_soil)

soilForest.f <- randomForest(response.soil~., data = rf.data.soil, ntree = 1000)
print(soilForest.f)
```


now predict treatment
```{r}

df_soilc<-as.data.frame(otu.soil)
taxa_names(Family.soil)
tax_table(Family.soil)
identical(colnames(df_soilc), taxa_names(Family.soil))
colnames(df_soilc)<-as.data.frame(tax_table(Family.soil))$Family
df_soilc


#predictors.soil=t(otu_table(Family.soil))
response.soilc <- as.factor(sample_data(Family.soil)$Treatment)
rf.data.soilc <- data.frame(response.soilc, df_soilc)

soilForest.fc <- randomForest(response.soilc~., data = rf.data.soilc, ntree = 1000)
print(soilForest.fc)
```

now just using the under samples 
```{r}
phy.still.under <- subset_samples(phy.still, Sampling.location=="U")

GPr.soilu  = transform_sample_counts(phy.still.under, function(x) x / sum(x) ) #transform samples based on relative abundance
GPr.soilu = filter_taxa(GPr.soilu, function(x) mean(x) > 1e-5, TRUE)

Family.soilu<-tax_glom(GPr.soilu,"Family")

otu.soilu<-t(otu_table(Family.soilu))

df_soilu<-as.data.frame(otu.soilu)
taxa_names(Family.soilu)
tax_table(Family.soilu)
identical(colnames(df_soilu), taxa_names(Family.soilu))
colnames(df_soilu)<-as.data.frame(tax_table(Family.soilu))$Family
df_soilu


predictors.soilu=t(otu_table(Family.soilu))
response.soilu <- as.factor(sample_data(Family.soilu)$Sampling.point)
rf.data.soilu <- data.frame(response.soilu, df_soilu)

soilForest.fu <- randomForest(response.soilu~., data = rf.data.soilu, ntree = 1000)
print(soilForest.fu)

```
saved 1100x650
```{r}
imp.soilu <- importance(soilForest.fu)#all the steps that are imp or imp. are building a dataframe that contains info about the taxa used by the Random Forest test to classify treatment 
imp.soilu <- data.frame(predictors = rownames(imp.soilu), imp.soilu)
imp.soilu.sort <- arrange(imp.soilu, dplyr::desc(MeanDecreaseGini))
imp.soilu.sort$predictors <- factor(imp.soilu.sort$predictors, levels = imp.soilu.sort$predictors)
imp.soilu.20 <- imp.soilu.sort[1:14, ]
imp.soilu.20

soilu_all_imp <- ggplot(`imp.soilu.20`, aes(x = predictors, y = MeanDecreaseGini)) +
  geom_bar(stat = "identity", fill = "lightblue4") +
  coord_flip() +
  xlab("Important Bacterial Families")+
  ylab("Mean Decrease Gini")+
  theme_bw()
soilu_all_imp


imp.soil.5<-imp.soilu.20 %>% head(7)
  
```

importance barplots
saved 1900x900
```{r, echo=TRUE}
imp.soil.5<-imp.soilu.20 %>% head(7)

names.fam<-factor(imp.soil.5$predictors)

soil.top.5<-rf.data.soilu %>% select(all_of(names.fam)) %>% mutate(Sampling.point = response.soilu)

soil.rf.barplot <- soil.top.5 %>% pivot_longer(1:7, names_to = "Family", values_to ="Abundance") %>% dplyr::mutate(Abundance = 100*Abundance) %>% 
dplyr::group_by(Sampling.point, Family) %>% dplyr::summarise(mean = mean(Abundance), 
                                            N = length(Abundance),
                 sd   = sd(Abundance),
                 se   = sd / sqrt(N)) %>% 
  ggplot(aes(x=Family,y=mean))+geom_bar(aes(fill = Family),colour="black", stat="identity")+ facet_grid(~Family)+xlab("Family")+ylab("Relative Abundance") + theme(axis.text.x = element_text(angle = 45, hjust = 1))+theme(axis.title.x=element_blank())+facet_wrap(~Sampling.point,scales = "free_y")+geom_errorbar(aes(ymin=mean-se,ymax=mean+se), width=0.5)+scale_fill_brewer(palette="Set1")

soil.rf.barplot
```


```{r, echo=TRUE}
cdataplots=ggplot(Trtdatas, aes(x=Family,y=mean))+geom_bar(aes(fill = Family),colour="black", stat="identity")+ facet_grid(~Family)+xlab("Family")+ylab("Relative Abundance") + theme(axis.text.x = element_text(angle = 45, hjust = 1))+theme(axis.title.x=element_blank())+facet_wrap(~Sampling.point,scales = "free_y")+geom_errorbar(aes(ymin=mean-se,ymax=mean+se))+scale_fill_brewer(palette="Set1")
cdataplots

cdataplots2=ggplot(Trtdatas, aes(x=Family,y=mean))+geom_bar(aes(fill = Family),colour="black", stat="identity")+ facet_grid(~Family)+ylab("Relative Abundance") +theme(axis.title.x=element_blank())+facet_wrap(~Sampling.point,scales = "free_y")+geom_errorbar(aes(ymin=mean-se,ymax=mean+se))+scale_fill_brewer(palette="Set1")
cdataplots2


compare_means(Abundance ~ sample_Family, data = df, group.by = "Family", p.adjust.method = "fdr",method="kruskal.test")
#bacterial relative abundance, kruskal-wallis
```


just control samples underneath carcass
```{r}
phy.still.under.cntrl <- subset_samples(phy.still.under, Treatment=="Control")

GPr.soiluc  = transform_sample_counts(phy.still.under.cntrl, function(x) x / sum(x) ) #transform samples based on relative abundance
GPr.soiluc = filter_taxa(GPr.soiluc, function(x) mean(x) > 1e-5, TRUE)

Family.soiluc<-tax_glom(GPr.soiluc,"Family")

otu.soiluc<-t(otu_table(Family.soiluc))

df_soiluc<-as.data.frame(otu.soiluc)
taxa_names(Family.soiluc)
tax_table(Family.soiluc)
identical(colnames(df_soiluc), taxa_names(Family.soiluc))
colnames(df_soiluc)<-as.data.frame(tax_table(Family.soiluc))$Family
df_soiluc


#predictors.soiluc=t(otu_table(Family.soiluc))
response.soiluc <- as.factor(sample_data(Family.soiluc)$Sampling.point)
rf.data.soiluc <- data.frame(response.soiluc, df_soiluc)

soilForest.fuc <- randomForest(response.soiluc~., data = rf.data.soiluc, ntree = 1000)
print(soilForest.fuc)
```

bleach samples underneath carcass now
```{r}
phy.still.under.bleach <- subset_samples(phy.still.under, Treatment=="Bleach")

GPr.soilub  = transform_sample_counts(phy.still.under.bleach, function(x) x / sum(x) ) #transform samples based on relative abundance
GPr.soilub = filter_taxa(GPr.soilub, function(x) mean(x) > 1e-5, TRUE)

Family.soilub<-tax_glom(GPr.soilub,"Family")

otu.soilub<-t(otu_table(Family.soilub))

df_soilub<-as.data.frame(otu.soilub)
taxa_names(Family.soilub)
tax_table(Family.soilub)
identical(colnames(df_soilub), taxa_names(Family.soilub))
colnames(df_soilub)<-as.data.frame(tax_table(Family.soilub))$Family
df_soilub


response.soilub <- as.factor(sample_data(Family.soilub)$Sampling.point)
rf.data.soilub <- data.frame(response.soilub, df_soilub)

soilForest.fub <- randomForest(response.soilub~., data = rf.data.soilub, ntree = 1000)
print(soilForest.fub)
```



***********************************
read in all beetle and swab data
```{r}
set.seed(2000)
setwd("C:/beetle_data_analysis/summer_beetles_all")
phy.all.summerba<-qza_to_phyloseq("sba-no-mito.qza","sba-rooted-tree.qza","sba-taxonomy.qza","summer_meta_allbeetles.txt")
```

now subset only beetles 
```{r}
phy.summer.beetles <- subset_samples(phy.all.summerba, Order=="Coleoptera")

```
only summer beetles now 
```{r}
phy.ch4.beetles1a <- subset_samples(phy.summer.beetles, Carcass %in% c("C22", "C23", "C24", "C25", "C26", "B1", "B2", "B3", "B4", "B5"))
phy.ch4.beetles2a <- subset_samples(phy.ch4.beetles1a, Family != "")#only beetles identified to family included

```


and only swabs 
```{r}
phy.summer.swabs <- subset_samples(phy.all.summerba, Order=="Swab")
phy.ch4.swabsa <- subset_samples(phy.summer.swabs, Carcass %in% c("C22", "C23", "C24", "C25", "C26", "B1", "B2", "B3", "B4", "B5"))

```

beetles and swabs
```{r}
phy.swab.beetle <- subset_samples(phy.all.summerba, Order %in% c("Coleoptera", "Swab"))
phy.swab.beetle <- subset_samples(phy.swab.beetle, Carcass %in% c("C22", "C23", "C24", "C25", "C26", "B1", "B2", "B3", "B4", "B5"))
```



beetles!
alpha diversity 
```{r}
beetle_richness_all2 <- estimate_richness(beetles_fam, measures = c("Observed", "Shannon", "InvSimpson"))
setwd("C:/soil_data/summer.soil")
beetle_meta <- read.delim("beetles_meta.txt") #metadata file with Sample.ID and sample names with . not -
beetle_richness4 <- setDT(beetle_richness_all2, keep.rownames = "Sample.ID")
beetle.div.meta <- merge(beetle_richness4, beetle_meta, by='Sample.ID', all.x=TRUE)
beetle.div.meta2 <- merge(beetle_richness4, beetle_meta, by='Sample.ID', all.x=TRUE)
file.pathb <- "C:/soil_data/summer.soil/plots/beetle_alpha_wmeta.csv"
write.csv(beetle.div.meta2, file=file.pathb)
```

pcoa first
saving 1050x700
```{r}
beetles_fam <- tax_glom(phy.ch4.beetles2a, taxrank = "Family")
ord.beetles = ordinate(phy.ch4.beetles2a, "PCoA", "unifrac", weighted=TRUE)
ord.beetles.fam = ordinate(beetles_fam, "PCoA", "unifrac", weighted=TRUE)

plot.ord.sbeetles <- plot_ordination(phy.ch4.beetles2a, ord.beetles, color = "Carcass", shape="Sampling.point" )+
  geom_point(alpha = 0.7,size = 8) +
  xlab("PCoA1 (54%)") +
  ylab("PCoA2 (32.2%)") +
  theme_classic() +
  scale_color_brewer (palette = "Set3")+
  scale_fill_discrete(labels = c("Control", "Under Carcass"))+
  stat_ellipse(level=0.95)
plot.ord.sbeetles$layers <- plot.ord.sbeetles$layers[-1]
plot.ord.sbeetles

plot.ord.sbeetles2 <- plot_ordination(phy.ch4.beetles2a, ord.beetles, color = "Family", shape="Sampling.point" )+
  geom_point(alpha = 0.7,size = 8) +
  xlab("PCoA1 (54%)") +
  ylab("PCoA2 (32.2%)") +
  theme_classic() +
  scale_color_brewer (palette = "Set3")+
  stat_ellipse(level=0.95)
plot.ord.sbeetles2$layers <- plot.ord.sbeetles2$layers[-1]
plot.ord.sbeetles2


##family level 
plot.ord.sbeetles.fam <- plot_ordination(phy.ch4.beetles2a, ord.beetles.fam,color = "Family", shape="Sampling.point" )+
  geom_point(alpha = 0.7,size = 8) +
  xlab("PCoA1 (23.1%)") +
  ylab("PCoA2 (13.9%)") +
  theme_classic() +
  scale_color_brewer (palette = "Set3")+
  stat_ellipse(level=0.95)
plot.ord.sbeetles.fam$layers <- plot.ord.sbeetles.fam$layers[-1]
plot.ord.sbeetles.fam

```


beetle bar plot of top 11 taxa 
```{r}
merge_less_than_top <- function(pobject, top=7){
  transformed <- transform_sample_counts(pobject, function(x) x/sum(x))
  otu.table <- as.data.frame(otu_table(transformed))
  otu.sort <- otu.table[order(rowMeans(otu.table), decreasing = TRUE),]
  otu.list <- row.names(otu.sort[(top+1):nrow(otu.sort),])
  merged <- merge_taxa(transformed, otu.list, 1)
  for (i in 1:dim(tax_table(merged))[1]){
    if (is.na(tax_table(merged)[i,2])){
      taxa_names(merged)[i] <- "Other"
      tax_table(merged)[i,1:7] <- "Other"}
  }
  return(merged)
}
family_orderb <- c("Enterobacteriaceae", "Erysipelotrichaceae", "Lactobacillaceae", "Morganellaceae", "Planococcaceae", "Rhizobiaceae", "Spiroplasmataceae", "Streptococcaceae", "Vagococcaceae", "Wohlfahrtiimonadaceae", "Xanthobacteraceae", "Other")
family_colorsb <- c("Enterobacteriaceae"="#8DD3C7", "Erysipelotrichaceae"="#FFFFB3", "Lactobacillaceae"="#BEBADA", "Morganellaceae"="#FB8072", "Planococcaceae"="#80B1D3", "Rhizobiaceae"="#FDB462", "Spiroplasmataceae"="#B3DE69", "Streptococcaceae"="#FCCDE5" , "Vagococcaceae"="#FFED6F", "Wohlfahrtiimonadaceae"="#BC80BD", "Xanthobacteraceae"="#CCEBC5", "Other"="#D9D9D9" )


beetle.fam.top10 <- merge_less_than_top(beetles_fam, top=9)

bnew2 <- plot_bar(beetle.fam.top10,fill = "Family")+
  scale_fill_brewer(palette="Set3")+
  ylab("Relative Abundance")
bnew2 <- bnew2 + facet_wrap(~sample_Family, scales="free_x") #gets rid of the empty space
bnew2

beetle.fam.top12 <- merge_less_than_top(beetles_fam, top=11)

bnew3 <- plot_bar(beetle.fam.top12,fill = "Family")+
  scale_fill_manual(values = family_colorsb, breaks = family_orderb) +
  ylab("Relative Abundance")
bnew3 <- bnew3 + facet_wrap(~sample_Family, scales="free_x") #gets rid of the empty space
bnew3

bnew4 <- plot_bar(beetle.fam.top12,fill = "Family")+
  scale_fill_manual(values = family_colorsb, breaks = family_orderb) +
  ylab("Relative Abundance (%)")
bnew4 <- bnew4 + facet_grid(~sample_Family, scales="free_x", space="free_x") #gets rid of the empty space
bnew4

bnew5 <- plot_bar(beetle.fam.top12,fill = "Family")+
  scale_fill_manual(values = family_colorsb, breaks = family_orderb) +
  ylab("Relative Abundance")
bnew5 <- bnew5 + facet_grid(~sample_Family, scales="free_x", space="free_x")+theme(
    strip.text.x = element_text(
      angle = 45,     # angle upwards
      hjust = 0,      # left-justify so text extends into the panel
      vjust = 0.5,
      size = 9
    )) #gets rid of the empty space
bnew5

```



beetle permanova
permanova
```{r}
GPdist.beetle=phyloseq::distance(phy.ch4.beetles2a, "wunifrac")

beetle.adonis <- adonis2(GPdist.beetle ~ Family*Sampling.point, as(sample_data(phy.ch4.beetles2a), "data.frame"), by="terms")
beetle.adonis

beetle.adonis2 <- adonis2(GPdist.beetle ~ Family*Carcass, as(sample_data(phy.ch4.beetles2a), "data.frame"), by="terms")
beetle.adonis2

```
family level permanova
```{r}
GPdist.beetlef=phyloseq::distance(beetles_fam, "wunifrac")

beetle.adonisf <- adonis2(GPdist.beetlef ~ Family*Sampling.point, as(sample_data(beetles_fam), "data.frame"), by="terms")
beetle.adonisf

beetle.adonis2f <- adonis2(GPdist.beetlef ~ Family*Carcass, as(sample_data(beetles_fam), "data.frame"), by="terms")
beetle.adonis2f

beetle.adonis2f2 <- adonis2(GPdist.beetlef ~ Family*Treatment.1, as(sample_data(beetles_fam), "data.frame"), by="terms")
beetle.adonis2f2
```



random forest
beetles 
```{r}
set.seed(2000)
GPr.beetles  = transform_sample_counts(phy.ch4.beetles2a, function(x) x / sum(x) ) #transform samples based on relative abundance
GPr.beetles = filter_taxa(GPr.beetles, function(x) mean(x) > 1e-5, TRUE)
Family.b<-tax_glom(GPr.beetles,"Family")

otu.famsb<-t(otu_table(Family.b))

df_famsb<-as.data.frame(otu.famsb)
taxa_names(Family.b)
tax_table(Family.b)
identical(colnames(df_famsb), taxa_names(Family.b))
colnames(df_famsb)<-as.data.frame(tax_table(Family.b))$Family
df_famsb


predictors.swabsb=t(otu_table(Family.b))
response.beetlesf <- as.factor(sample_data(Family.b)$Family)
rf.data.beetlesf <- data.frame(response.beetlesf, df_famsb)

betForest.f <- randomForest(response.beetlesf~., data = rf.data.beetlesf, ntree = 1000)
print(betForest.f)#returns overall Random Forest results
```
saved 1100x650
```{r}
imp.bet <- importance(betForest.f)#all the steps that are imp or imp. are building a dataframe that contains info about the taxa used by the Random Forest test to classify treatment 
imp.bet <- data.frame(predictors = rownames(imp.bet), imp.bet)
imp.bet.sort <- arrange(imp.bet, dplyr::desc(MeanDecreaseGini))
imp.bet.sort$predictors <- factor(imp.bet.sort$predictors, levels = imp.bet.sort$predictors)
imp.bet.20 <- imp.bet.sort[1:14, ]
imp.bet.20

bet_all_imp <- ggplot(`imp.bet.20`, aes(x = predictors, y = MeanDecreaseGini)) +
  geom_bar(stat = "identity", fill = "lightblue4") +
  coord_flip() +
  xlab("Important Bacterial Families")+
  ylab("Mean Decrease Gini")+
  theme_bw()
bet_all_imp
  
```

saved 1900x900
```{r, echo=TRUE}
imp.bet.5<-imp.bet.20 %>% head(7)

names.fam<-factor(imp.bet.5$predictors)

bet.top.5<-rf.data.beetlesf %>% select(all_of(names.fam)) %>% mutate(Family = response.beetlesf)

bet.rf.barplot <- bet.top.5 %>% pivot_longer(1:7, names_to = "Family.tax", values_to ="Abundance") %>% dplyr::mutate(Abundance = 100*Abundance) %>% 
dplyr::group_by( Family,Family.tax) %>% dplyr::summarise(mean = mean(Abundance), 
                                            N = length(Abundance),
                 sd   = sd(Abundance),
                 se   = sd / sqrt(N)) %>% 
  ggplot(aes(x=Family.tax,y=mean))+geom_bar(aes(fill = Family.tax),colour="black", stat="identity")+ facet_grid(~Family)+xlab("Family")+ylab("Relative Abundance") + theme(axis.text.x = element_text(angle = 45, hjust = 1))+theme(axis.title.x=element_blank())+facet_wrap(~Family,scales = "free_y")+geom_errorbar(aes(ymin=mean-se,ymax=mean+se), width=0.5)+scale_fill_brewer(palette="Set1")

bet.rf.barplot
```


beetle internal microbiome random forest, bleach and control carcasses separately 
```{r}
#first bleach

phy.beetles.bleach <- subset_samples(phy.ch4.beetles2a, Carcass %in% c("B1", "B2", "B3", "B4", "B5"))

#beetles_bleach_fam <- tax_glom(phy.beetles.bleach, taxrank = "Family")

set.seed(2000)
GPr.beetlesb  = transform_sample_counts(phy.beetles.bleach, function(x) x / sum(x) ) #transform samples based on relative abundance
GPr.beetlesb = filter_taxa(GPr.beetlesb, function(x) mean(x) > 1e-5, TRUE)

Family.bb<-tax_glom(GPr.beetlesb,"Family")

otu.famsbb<-t(otu_table(Family.bb))

df_famsbb<-as.data.frame(otu.famsbb)
taxa_names(Family.bb)
tax_table(Family.bb)
identical(colnames(df_famsbb), taxa_names(Family.bb))
colnames(df_famsbb)<-as.data.frame(tax_table(Family.bb))$Family
df_famsbb


predictors.swabsbb=t(otu_table(Family.bb))
response.beetlesfb <- as.factor(sample_data(Family.bb)$Family)
rf.data.beetlesfb <- data.frame(response.beetlesfb, df_famsbb)

betForest.fb <- randomForest(response.beetlesfb~., data = rf.data.beetlesfb, ntree = 1000)
print(betForest.fb)#returns overall Random Forest results

```
bleach importance plots
```{r}
imp.betb <- importance(betForest.fb)#all the steps that are imp or imp. are building a dataframe that contains info about the taxa used by the Random Forest test to classify treatment 
imp.betb <- data.frame(predictors = rownames(imp.betb), imp.betb)
imp.sort.betb <- arrange(imp.betb, dplyr::desc(MeanDecreaseGini))
imp.sort.betb$predictors <- factor(imp.sort.betb$predictors, levels = imp.sort.betb$predictors)
imp.20.betb <- imp.sort.betb[1:15, ]
imp.bet.plotb <- ggplot(imp.20.betb, aes(x = predictors, y = MeanDecreaseGini)) +
  geom_bar(stat = "identity", fill = "lightblue4") +
  coord_flip() +
  ggtitle("Most important families bleach beetles")#\n in a string tells it to start a new line
imp.bet.plotb
#imp.20$MeanDecreaseGini
otunames.betb <- imp.20.betb$predictors
r.betb <- rownames(tax_table(Family.betb)) %in% otunames.betb
otunames.betb
PredictorTable.betb<-kable(tax_table(Family.betb)[r.betb, ])#returns a list of the most important predictors for Random Forest Classification
PredictorTable.betb
```

importance barplots
saved 1900x900
```{r, echo=TRUE}
FambsRandomForestSubsetb = subset_taxa(Family.betb, row.names(tax_table(Family.betb))%in% otunames.betb)
FambsRandomForestSubsetb

dfb <- psmelt(FambsRandomForestSubsetb)
dfb$Abundance=dfb$Abundance*100
Trtdatab <- ddply(dfb, c("sample_Family", "Family"), summarise,
                 N    = length(Abundance),
                 mean = mean(Abundance),
                 sd   = sd(Abundance),
                 se   = sd / sqrt(N)
)
Trtdatab

cdataplotb=ggplot(Trtdatab, aes(x=Family,y=mean))+geom_bar(aes(fill = Family),colour="black", stat="identity")+ facet_grid(~Family)+xlab("Family")+ylab("Relative Abundance") + theme(axis.text.x = element_text(angle = 45, hjust = 1))+theme(axis.title.x=element_blank())+facet_wrap(~sample_Family,scales = "free_y")+geom_errorbar(aes(ymin=mean-se,ymax=mean+se))+scale_fill_brewer(palette="Set1")
cdataplotb

cdataplot2b=ggplot(Trtdatab, aes(x=Family,y=mean))+geom_bar(aes(fill = Family),colour="black", stat="identity")+ facet_grid(~Family)+ylab("Relative Abundance") +theme(axis.title.x=element_blank())+facet_wrap(~sample_Family,scales = "free_y")+geom_errorbar(aes(ymin=mean-se,ymax=mean+se))+scale_fill_brewer(palette="Set1")
cdataplot2b


compare_means(Abundance ~ sample_Family, data = dfb, group.by = "Family", p.adjust.method = "fdr",method="kruskal.test")
#bacterial relative abundance, kruskal-wallis
```

now control carcasses only 
```{r}
phy.beetles.cntrl <- subset_samples(phy.ch4.beetles2a, Carcass %in% c("C22", "C23", "C24", "C25", "C26"))

set.seed(2000)
GPr.beetlesc  = transform_sample_counts(phy.beetles.cntrl, function(x) x / sum(x) ) #transform samples based on relative abundance
GPr.beetlesc = filter_taxa(GPr.beetlesc, function(x) mean(x) > 1e-5, TRUE)

Family.bbc<-tax_glom(GPr.beetlesc,"Family")

otu.famsbbc<-t(otu_table(Family.bbc))

df_famsbbc<-as.data.frame(otu.famsbbc)
taxa_names(Family.bbc)
tax_table(Family.bbc)
identical(colnames(df_famsbbc), taxa_names(Family.bbc))
colnames(df_famsbbc)<-as.data.frame(tax_table(Family.bbc))$Family
df_famsbbc


predictors.swabsbbc=t(otu_table(Family.bbc))
response.beetlesfbc <- as.factor(sample_data(Family.bbc)$Family)
rf.data.beetlesfbc <- data.frame(response.beetlesfbc, df_famsbbc)

betForest.fc <- randomForest(response.beetlesfbc~., data = rf.data.beetlesfbc, ntree = 1000)
print(betForest.fc)

```

control importance plots
```{r}
imp.betc <- importance(betForest.fc)#all the steps that are imp or imp. are building a dataframe that contains info about the taxa used by the Random Forest testto classify treatment 
imp.betc <- data.frame(predictors = rownames(imp.betc), imp.betc)
imp.sort.betc <- arrange(imp.betc, dplyr::desc(MeanDecreaseGini))
imp.sort.betc$predictors <- factor(imp.sort.betc$predictors, levels = imp.sort.betc$predictors)
imp.20.betc <- imp.sort.betc[1:15, ]
imp.bet.plotc <- ggplot(imp.20.betc, aes(x = predictors, y = MeanDecreaseGini)) +
  geom_bar(stat = "identity", fill = "lightblue4") +
  coord_flip() +
  ggtitle("Most important families bleach beetles")#\n in a string tells it to start a new line
imp.bet.plotc
#imp.20$MeanDecreaseGini
otunames.betc <- imp.20.betc$predictors
r.betc <- rownames(tax_table(Family.betc)) %in% otunames.betc
otunames.betc
PredictorTable.betc<-kable(tax_table(Family.betc)[r.betc, ])#returns a list of the most important predictors for Random Forest Classification
PredictorTable.betc
```


importance plots 
```{r, echo=TRUE}
FambsRandomForestSubsetc = subset_taxa(Family.betc, row.names(tax_table(Family.betc))%in% otunames.betc)
FambsRandomForestSubsetc

dfc <- psmelt(FambsRandomForestSubsetc)
dfc$Abundance=dfc$Abundance*100
Trtdatac <- ddply(dfc, c("sample_Family", "Family"), summarise,
                 N    = length(Abundance),
                 mean = mean(Abundance),
                 sd   = sd(Abundance),
                 se   = sd / sqrt(N)
)
Trtdatac

cdataplotc=ggplot(Trtdatac, aes(x=Family,y=mean))+geom_bar(aes(fill = Family),colour="black", stat="identity")+ facet_grid(~Family)+xlab("Family")+ylab("Relative Abundance") + theme(axis.text.x = element_text(angle = 45, hjust = 1))+theme(axis.title.x=element_blank())+facet_wrap(~sample_Family,scales = "free_y")+geom_errorbar(aes(ymin=mean-se,ymax=mean+se))+scale_fill_brewer(palette="Set1")
cdataplotc


compare_means(Abundance ~ sample_Family, data = dfc, group.by = "Family", p.adjust.method = "fdr",method="kruskal.test")
```


swabs!
alpha diversity
```{r}
swab_richness_all2 <- estimate_richness(swabs_fam, measures = c("Observed", "Shannon", "InvSimpson"))
setwd("C:/soil_data/summer.soil")
swab_meta <- read.delim("swabs_meta.txt") #metadata file with Sample.ID and sample names with . not -
swab_richness4 <- setDT(swab_richness_all2, keep.rownames = "Sample.ID")
swab.div.meta <- merge(swab_richness4, swab_meta, by='Sample.ID', all.x=TRUE)
swab.div.meta2 <- merge(swab_richness4, swab_meta, by='Sample.ID', all.x=TRUE)
file.paths <- "C:/soil_data/summer.soil/plots/swab_alpha_wmeta3.csv"
write.csv(swab.div.meta2, file=file.paths)
```

anova for alpha diversity
```{r}
##splitting data frame to each body site 

#swabnew <- split(swab.div.meta2, swab.div.meta2$Family)
#mouthnew <- swabnew$Mouth
#recnew <- swabnew$Rectum
#sknew <- swabnew$Skin
#mouthnew<- mouthnew %>% mutate(Sampling.point = factor(Sampling.point))
#recnew<- recnew %>% mutate(Sampling.point = factor(Sampling.point))
#sknew<- sknew %>% mutate(Sampling.point = factor(Sampling.point))

mouth_alphao <- lm(Observed~Treatment*Sampling.point, data = mouthnew)
summary(mouth_alphao)
plot(mouth_alphao)

mouth_alphash <- lm(Shannon~Treatment*Sampling.point, data = mouthnew)
summary(mouth_alphash)
plot(mouth_alphash)

mouth_alphais <- lm(InvSimpson~Treatment*Sampling.point, data = mouthnew)
summary(mouth_alphais)
plot(mouth_alphais)

rec_alphao <- lm(Observed~Treatment*Sampling.point, data = recnew)
summary(rec_alphao)
plot(rec_alphao)

rec_alphas <- lm(Shannon~Treatment*Sampling.point, data = recnew)
summary(rec_alphas)
plot(rec_alphas)

rec_alphais <- lm(InvSimpson~Treatment*Sampling.point, data = recnew)
summary(rec_alphais)
plot(rec_alphais)

sk_alphao <- lm(Observed~Treatment*Sampling.point, data = sknew)
summary(sk_alphao)
plot(sk_alphao)

sk_alphas <- lm(Shannon~Treatment*Sampling.point, data = sknew)
summary(sk_alphas)
plot(sk_alphas)

sk_alphais <- lm(InvSimpson~Treatment*Sampling.point, data = sknew)
summary(sk_alphais)
plot(sk_alphais)
```
pot-hoc testing 
```{r}
contsko<-emmeans(sk_alphao,~Sampling.point*Treatment)
contsko
contrasko<-as.data.frame(contrast(contsko, method = "pairwise")) 
contrasko %>% arrange(p.value)

contsks<-emmeans(sk_alphas,~Sampling.point*Treatment)
contsks
contrasks<-as.data.frame(contrast(contsks, method = "pairwise")) 
contrasks %>% arrange(p.value)

contskis<-emmeans(sk_alphais,~Sampling.point*Treatment)
contskis
contraskis<-as.data.frame(contrast(contskis, method = "pairwise")) 
contraskis %>% arrange(p.value)

contmo<-emmeans(mouth_alphao,~Sampling.point*Treatment)
contmo
contramo<-as.data.frame(contrast(contmo, method = "pairwise")) 
contramo %>% arrange(p.value)

contms<-emmeans(mouth_alphash,~Sampling.point*Treatment)
contms
contrams<-as.data.frame(contrast(contms, method = "pairwise")) 
contrams %>% arrange(p.value)
```


permanova
testing for relationship between carcass body area and time point
```{r}
GPdist.swabsf=phyloseq::distance(swabs_fam, "wunifrac")

swabs.adonisf <- adonis2(GPdist.swabsf ~ Family*Sampling.point, as(sample_data(swabs_fam), "data.frame"), by="terms")
swabs.adonisf

swabs.adonis2f <- adonis2(GPdist.swabsf ~ Family*Carcass, as(sample_data(swabs_fam), "data.frame"), by="terms")
swabs.adonis2f
```
permanova of body sites comparing control and bleach
```{r}
swab_skin <- subset_samples(swabs_fam, Family=="Skin")
swab_mouth <- subset_samples(swabs_fam, Family=="Mouth")
swab_rec <- subset_samples(swabs_fam, Family=="Rectum")
```
```{r}
GPdist.skin=phyloseq::distance(swab_skin, "wunifrac")

swabs.adonissk <- adonis2(GPdist.skin ~ Carcass*Sampling.point, as(sample_data(swab_skin), "data.frame"), by="terms")
swabs.adonissk
```
```{r}
GPdist.rec=phyloseq::distance(swab_rec, "wunifrac")

swabs.adonisr <- adonis2(GPdist.rec ~ Carcass*Sampling.point, as(sample_data(swab_recci), "data.frame"), by="terms")
swabs.adonisr
```

pcoa
```{r}
ord.swabs = ordinate(phy.ch4.swabsa, "PCoA", "unifrac", weighted=TRUE)

plot.ord.sswab <- plot_ordination(phy.ch4.swabsa, ord.swabs, color = "Carcass", shape="Sampling.point" )+
  geom_point(alpha = 0.7,size = 8) +
  #xlab("PCoA1 (54%)") +
  #ylab("PCoA2 (32.2%)") +
  theme_classic() +
  scale_color_brewer (palette = "Set3")+
  scale_fill_discrete(labels = c("Control", "Under Carcass"))+
  stat_ellipse(level=0.95)
plot.ord.sswab$layers <- plot.ord.sswab$layers[-1]
plot.ord.sswab

plot.ord.sswab2 <- plot_ordination(phy.ch4.swabsa, ord.swabs, color = "Sampling.point", shape="Carcass" )+
  geom_point(alpha = 0.7,size = 8) +
  #xlab("PCoA1 (54%)") +
  #ylab("PCoA2 (32.2%)") +
  theme_classic() +
  scale_color_brewer (palette = "Set3")+
  scale_fill_discrete(labels = c("Control", "Under Carcass"))+
  stat_ellipse(level=0.95)
plot.ord.sswab2$layers <- plot.ord.sswab2$layers[-1]
plot.ord.sswab2
```

looking at individual body sites now
```{r}
plot.ord.sswabbs <- plot_ordination(phy.ch4.swabsa, ord.swabs, color = "Sampling.point", shape="Family" )+
  geom_point(alpha = 0.7,size = 8) +
  #xlab("PCoA1 (54%)") +
  #ylab("PCoA2 (32.2%)") +
  theme_classic() +
  scale_color_brewer (palette = "Set3")+
  stat_ellipse(level=0.95)
plot.ord.sswabbs$layers <- plot.ord.sswabbs$layers[-1]
plot.ord.sswabbs

plot.ord.sswabbs2 <- plot_ordination(phy.ch4.swabsa, ord.swabs, color = "Family", shape="Sampling.point" )+
  geom_point(alpha = 0.7,size = 8) +
  #xlab("PCoA1 (54%)") +
  #ylab("PCoA2 (32.2%)") +
  theme_classic() +
  scale_color_brewer (palette = "Set3")+
  stat_ellipse(level=0.95)
plot.ord.sswabbs2$layers <- plot.ord.sswabbs2$layers[-1]
plot.ord.sswabbs2
```

swabs family level 
```{r}

ord.swabs.fam = ordinate(swabs_fam, "PCoA", "unifrac", weighted=TRUE)

plot.ord.sswab.fam <- plot_ordination(swabs_fam, ord.swabs.fam, color = "Carcass", shape="Sampling.point" )+
  geom_point(alpha = 0.7,size = 8) +
  #xlab("PCoA1 (54%)") +
  #ylab("PCoA2 (32.2%)") +
  theme_classic() +
  scale_color_brewer (palette = "Set3")+
  scale_fill_discrete(labels = c("Control", "Under Carcass"))+
  stat_ellipse(level=0.95)
plot.ord.sswab.fam$layers <- plot.ord.sswab.fam$layers[-1]
plot.ord.sswab.fam

plot.ord.sswab.fam2 <- plot_ordination(swabs_fam, ord.swabs.fam, color = "Carcass",shape="Family" )+
  geom_point(alpha = 0.7,size = 8) +
  xlab("PCoA1 (37.3%)") +
  ylab("PCoA2 (16.7%)") +
  theme_classic() +
  scale_color_brewer (palette = "Set3")+
  scale_fill_discrete(labels = c("Control", "Under Carcass"))+
  stat_ellipse(level=0.95)
plot.ord.sswab.fam2$layers <- plot.ord.sswab.fam2$layers[-1]
plot.ord.sswab.fam2

plot.ord.sswab.fam3 <- plot_ordination(swabs_fam, ord.swabs.fam, color = "Sampling.point",shape="Family" )+
  geom_point(alpha = 0.7,size = 8) +
  xlab("PCoA1 (37.3%)") +
  ylab("PCoA2 (16.7%)") +
  theme_classic() +
  scale_color_brewer (palette = "Set3")+
  guides(color=guide_legend(title = "Sampling point"))+
  stat_ellipse(level=0.95)
plot.ord.sswab.fam3$layers <- plot.ord.sswab.fam3$layers[-1]
plot.ord.sswab.fam3

```


swabs bar plot
saved 2050 x 800 (i think)
```{r}
merge_less_than_top <- function(pobject, top=7){
  transformed <- transform_sample_counts(pobject, function(x) x/sum(x))
  otu.table <- as.data.frame(otu_table(transformed))
  otu.sort <- otu.table[order(rowMeans(otu.table), decreasing = TRUE),]
  otu.list <- row.names(otu.sort[(top+1):nrow(otu.sort),])
  merged <- merge_taxa(transformed, otu.list, 1)
  for (i in 1:dim(tax_table(merged))[1]){
    if (is.na(tax_table(merged)[i,2])){
      taxa_names(merged)[i] <- "Other"
      tax_table(merged)[i,1:7] <- "Other"}
  }
  return(merged)
}

family_orders <- c("Clostridiaceae", "Enterobacteriaceae", "Enterococcaceae", "Lactobacillaceae", "Moraxellaceae", "Morganellaceae", "Planococcaceae","Staphylococcaceae", "Streptococcaceae", "Vagococcaceae", "Wohlfahrtiimonadaceae", "Other")
family_colorss <- c("Clostridiaceae"="#8DD3C7", "Enterobacteriaceae"="#FFFFB3", "Enterococcaceae"="#BEBADA", "Lactobacillaceae"="#FB8072", "Moraxellaceae"="#80B1D3", "Morganellaceae"="#FDB462", "Planococcaceae"="#B3DE69", "Staphylococcaceae"="#FCCDE5" , "Streptococcaceae"="#FFED6F", "Vagococcaceae"="#BC80BD", "Wohlfahrtiimonadaceae"="#CCEBC5", "Other"="#D9D9D9" )


swabs_fam <- tax_glom(phy.ch4.swabsa, taxrank = "Family")
swab.fam.top10 <- merge_less_than_top(swabs_fam, top=9)

snew2 <- plot_bar(swab.fam.top10,fill = "Family")+
  scale_fill_brewer(palette="Set3")+
  ylab("Relative Abundance")
snew2 <- snew2 + facet_wrap(~sample_Family, scales="free_x") #gets rid of the empty space
snew2

swab.fam.top12 <- merge_less_than_top(swabs_fam, top=11)

snew3 <- plot_bar(swab.fam.top12,fill = "Family")+
  scale_fill_manual(values = family_colorss, breaks = family_orders)+
  ylab("Relative Abundance (%)")
snew3 <- snew3 + facet_wrap(~sample_Family, scales="free_x") #gets rid of the empty space
snew3


snew4 <- plot_bar(swab.fam.top12, fill = "Family") +
  scale_fill_manual(values = family_colorss, breaks = family_orders)+
  ylab("Relative Abundance (%)")+
  facet_grid(. ~ sample_Family + Sampling.point, scales = "free_x")

snew4
```

permanova by individual body sites 
comparing treatment and time
```{r}
swab_skin <- subset_samples(swabs_fam, Family=="Skin")
swab_mouth <- subset_samples(swabs_fam, Family=="Mouth")
swab_rec <- subset_samples(swabs_fam, Family=="Rectum")

GPdist.skin=phyloseq::distance(swab_skin, "wunifrac")


swabs.adonissk2 <- adonis2(GPdist.skin ~ Carcass, as(sample_data(swab_skin), "data.frame"), by="terms")
swabs.adonissk2
swabs.adonissk3 <- adonis2(GPdist.skin ~ Sampling.point, as(sample_data(swab_skin), "data.frame"), by="terms")
swabs.adonissk3

GPdist.rec=phyloseq::distance(swab_rec, "wunifrac")

swabs.adonisr <- adonis2(GPdist.rec ~ Carcass*Sampling.point, as(sample_data(swab_rec), "data.frame"), by="terms")
swabs.adonisr

GPdist.mo=phyloseq::distance(swab_mouth, "wunifrac")

swabs.adonism <- adonis2(GPdist.mo ~ Carcass*Sampling.point, as(sample_data(swab_mouth), "data.frame"), by="terms")
swabs.adonism
```
```{r}
swabs.adonissk2 <- adonis2(GPdist.skin ~ Carcass, as(sample_data(swab_skin), "data.frame"), by="terms")
swabs.adonissk2
swabs.adonissk3 <- adonis2(GPdist.skin ~ Sampling.point, as(sample_data(swab_skin), "data.frame"), by="terms")
swabs.adonissk3
swabs.adonisr2 <- adonis2(GPdist.rec ~ Carcass, as(sample_data(swab_rec), "data.frame"), by="terms")
swabs.adonisr2
swabs.adonisr3 <- adonis2(GPdist.rec ~ Sampling.point, as(sample_data(swab_rec), "data.frame"), by="terms")
swabs.adonisr3
swabs.adonism2 <- adonis2(GPdist.mo ~ Carcass, as(sample_data(swab_mouth), "data.frame"), by="terms")
swabs.adonism2
swabs.adonism3 <- adonis2(GPdist.mo ~ Sampling.point, as(sample_data(swab_mouth), "data.frame"), by="terms")
swabs.adonism3
```



swab random forest
together, then by body site 
testing for time
```{r}
#beetles_bleach_fam <- tax_glom(phy.beetles.bleach, taxrank = "Family")

set.seed(2000)
GPr.swabs  = transform_sample_counts(phy.ch4.swabsa, function(x) x / sum(x) ) #transform samples based on relative abundance
GPr.swabs = filter_taxa(GPr.swabs, function(x) mean(x) > 1e-5, TRUE)

Family.swab<-tax_glom(GPr.swabs,"Family")


otu.fam<-t(otu_table(Family.swab))

df_fam<-as.data.frame(otu.fam)
taxa_names(Family.swab)
tax_table(Family.swab)
identical(colnames(df_fam), taxa_names(Family.swab))
colnames(df_fam)<-as.data.frame(tax_table(Family.swab))$Family
df_fam

Family.swab<-df_fam
ps <- name_na_taxa(Family.swab)
ps <- label_duplicate_taxa(Family.swab, tax_level = "Family", asv_as_id = F)


family_counts_table <- as.data.frame(otu_table(ps))

as.data.table(t(otu_table(Family.swab)))


predictors.swab=t(otu_table(Family.swab))
response.swab <- as.factor(sample_data(Family.swab)$Sampling.point)
rf.data.swab <- data.frame(response.swab, df_fam)

swabForest.f <- randomForest(response.swab~., data = rf.data.swab, ntree = 1000)
print(swabForest.f)#returns overall Random Forest results

```

saved 1100x650
```{r}
imp.af <- importance(swabForest.f)#all the steps that are imp or imp. are building a dataframe that contains info about the taxa used by the Random Forest test to classify treatment 
imp.af <- data.frame(predictors = rownames(imp.af), imp.af)
imp.af.sort <- arrange(imp.af, dplyr::desc(MeanDecreaseGini))
imp.af.sort$predictors <- factor(imp.af.sort$predictors, levels = imp.af.sort$predictors)
imp.af.20 <- imp.af.sort[1:14, ]
imp.af.20

swab_all_imp <- ggplot(`imp.af.20`, aes(x = predictors, y = MeanDecreaseGini)) +
  geom_bar(stat = "identity", fill = "lightblue4") +
  coord_flip() +
  xlab("Important Bacterial Families")+
  ylab("Mean Decrease Gini")+
  theme_bw()
swab_all_imp
  
```


saved 1900x900
```{r, echo=TRUE}
imp.af.5<-imp.af.20 %>% head(7)

names.fam<-factor(imp.af.5$predictors)

af.top.5<-rf.data.swab %>% select(all_of(names.fam)) %>% mutate(Sampling.point = response.swab)

af.rf.barplot <- af.top.5 %>% pivot_longer(1:7, names_to = "Family", values_to ="Abundance") %>% dplyr::mutate(Abundance = 100*Abundance) %>% 
dplyr::group_by(Sampling.point,Family) %>% dplyr::summarise(mean = mean(Abundance), 
                                            N = length(Abundance),
                 sd   = sd(Abundance),
                 se   = sd / sqrt(N)) %>% 
  ggplot(aes(x=Family,y=mean))+geom_bar(aes(fill = Family),colour="black", stat="identity")+ facet_grid(~Family)+xlab("Family")+ylab("Relative Abundance") + theme(axis.text.x = element_text(angle = 45, hjust = 1))+theme(axis.title.x=element_blank())+facet_wrap(~Sampling.point,scales = "free_y")+geom_errorbar(aes(ymin=mean-se,ymax=mean+se), width=0.5)+scale_fill_brewer(palette="Set1")

af.rf.barplot
```


testing for body site 
```{r}
response.swabnew <- as.factor(sample_data(Family.swab)$Family)
rf.data.swab.new <- data.frame(response.swabnew, predictors.swab)

swabForest.f2 <- randomForest(response.swabnew~., data = rf.data.swab.new, ntree = 1000)
print(swabForest.f2)#returns overall Random Forest results

```

now splitting by body site 
```{r}
phy.swabs.mouth <- subset_samples(phy.ch4.swabsa, Family=="Mouth")

set.seed(2000)
GPr.swabsm  = transform_sample_counts(phy.swabs.mouth, function(x) x / sum(x) ) #transform samples based on relative abundance
GPr.swabsm = filter_taxa(GPr.swabsm, function(x) mean(x) > 1e-5, TRUE)

Family.swabm<-tax_glom(GPr.swabsm,"Family")

otu.famm<-t(otu_table(Family.swabm))

df_famm<-as.data.frame(otu.famm)
taxa_names(Family.swabm)
tax_table(Family.swabm)
identical(colnames(df_famm), taxa_names(Family.swabm))
colnames(df_famm)<-as.data.frame(tax_table(Family.swabm))$Family
df_famm


predictors.swabm=t(otu_table(Family.swabm))
response.swabm <- as.factor(sample_data(Family.swabm)$Sampling.point)
rf.data.swabm <- data.frame(response.swabm, df_famm)

swabForest.fm <- randomForest(response.swabm~., data = rf.data.swabm, ntree = 1000)
print(swabForest.fm)

```

skin only 
```{r}
phy.swabs.skin <- subset_samples(phy.ch4.swabsa, Family=="Skin")

set.seed(2000)
GPr.swabss  = transform_sample_counts(phy.swabs.skin, function(x) x / sum(x) ) #transform samples based on relative abundance
GPr.swabss = filter_taxa(GPr.swabss, function(x) mean(x) > 1e-5, TRUE)

Family.swabs<-tax_glom(GPr.swabss,"Family")

otu.fams<-t(otu_table(Family.swabs))

df_fams<-as.data.frame(otu.fams)
taxa_names(Family.swabs)
tax_table(Family.swabs)
identical(colnames(df_fams), taxa_names(Family.swabs))
colnames(df_fams)<-as.data.frame(tax_table(Family.swabs))$Family
df_fams


predictors.swabs=t(otu_table(Family.swabs))
response.swabs <- as.factor(sample_data(Family.swabs)$Sampling.point)
rf.data.swabs <- data.frame(response.swabs, df_fams)

swabForest.fs <- randomForest(response.swabs~., data = rf.data.swabs, ntree = 1000)
print(swabForest.fs)
```
rectum only
```{r}
phy.swabs.rectum <- subset_samples(phy.ch4.swabsa, Family=="Rectum")

set.seed(2000)
GPr.swabr  = transform_sample_counts(phy.swabs.rectum, function(x) x / sum(x) ) #transform samples based on relative abundance
GPr.swabr = filter_taxa(GPr.swabr, function(x) mean(x) > 1e-5, TRUE)

Family.swabsr<-tax_glom(GPr.swabr,"Family")

otu.famsr<-t(otu_table(Family.swabsr))

df_famsr<-as.data.frame(otu.famsr)
taxa_names(Family.swabsr)
tax_table(Family.swabsr)
identical(colnames(df_famsr), taxa_names(Family.swabsr))
colnames(df_famsr)<-as.data.frame(tax_table(Family.swabsr))$Family
df_famsr


predictors.swabsr=t(otu_table(Family.swabsr))
response.swabsr <- as.factor(sample_data(Family.swabsr)$Sampling.point)
rf.data.swabsr <- data.frame(response.swabsr, df_famsr)

swabForest.fr <- randomForest(response.swabsr~., data = rf.data.swabsr, ntree = 1000)
print(swabForest.fr)
```

all swabs and time importance plots
```{r}
imp.swab <- importance(swabForest.f)#all the steps that are imp or imp. are building a dataframe that contains info about the taxa used by the Random Forest testto classify treatment 
imp.swab <- data.frame(predictors = rownames(imp.swab), imp.swab)
imp.sort.swab <- arrange(imp.swab, dplyr::desc(MeanDecreaseGini))
imp.sort.swab$predictors <- factor(imp.sort.swab$predictors, levels = imp.sort.swab$predictors)
imp.20.swab <- imp.sort.swab[1:15, ]
imp.bet.plotsa <- ggplot(imp.20.swab, aes(x = predictors, y = MeanDecreaseGini)) +
  geom_bar(stat = "identity", fill = "lightblue4") +
  coord_flip() +
  ggtitle("Most important families bleach beetles")#\n in a string tells it to start a new line
imp.bet.plotsa
#imp.20$MeanDecreaseGini
otunames.swab <- imp.20.swab$predictors
r.swab <- rownames(tax_table(Family.swab)) %in% otunames.swab
otunames.swab
PredictorTable.swab<-kable(tax_table(Family.swab)[r.swab, ])#returns a list of the most important predictors for Random Forest Classification
PredictorTable.swab
```

importance barplots
saved 1900x900
```{r, echo=TRUE}
FambsRandomForestSubsetfa = subset_taxa(Family.swab, row.names(tax_table(Family.swab))%in% otunames.swab)
FambsRandomForestSubsetfa

dfsa <- psmelt(FambsRandomForestSubsetfa)
dfsa$Abundance=dfsa$Abundance*100
Trtdatasa <- ddply(dfsa, c("Sampling.point", "Family"), summarise,
                 N    = length(Abundance),
                 mean = mean(Abundance),
                 sd   = sd(Abundance),
                 se   = sd / sqrt(N)
)
Trtdatasa

cdataplotsa=ggplot(Trtdatasa, aes(x=Family,y=mean))+geom_bar(aes(fill = Family),colour="black", stat="identity")+ facet_grid(~Sampling.point)+xlab("Family")+ylab("Relative Abundance") + theme(axis.text.x = element_text(angle = 45, hjust = 1))+theme(axis.title.x=element_blank())+facet_wrap(~Sampling.point,scales = "free_y")+geom_errorbar(aes(ymin=mean-se,ymax=mean+se))+scale_fill_brewer(palette="Set1")
cdataplotsa

cdataplot2sa=ggplot(Trtdatasa, aes(x=Family,y=mean))+geom_bar(aes(fill = Family),colour="black", stat="identity")+ facet_grid(~Family)+ylab("Relative Abundance") +theme(axis.title.x=element_blank())+facet_wrap(~sample_Family,scales = "free_y")+geom_errorbar(aes(ymin=mean-se,ymax=mean+se))+scale_fill_brewer(palette="Set1")
cdataplot2sa


compare_means(Abundance ~ sample_Family, data = dfsa, group.by = "Family", p.adjust.method = "fdr",method="kruskal.test")
```



read in data of all three sample types (beetles, swabs, soil) + filter
```{r}
setwd("C:/soil_data/summer.soil")

#all together
phy.summer4 <-qza_to_phyloseq(features="summer-table-no-mito.qza",taxonomy="summer-taxonomy.qza",metadata="summer-meta2.tsv")

phy.summer.filt4 <- prune_samples(sample_sums(phy.summer4) >= 12000, phy.summer4) #removing samples that have less than 12000 reads, not rarefying any 
phy.stillnew2 <- subset_samples(phy.summer.filt4, Size=="Still")
ch4all <- merge_phyloseq(phy.stillnew2, phy.swab.beetle5)
ch4all <- subset_samples(ch4all, Family != "")

```
treatments
```{r}
ch4bleach <- subset_samples(ch4all, Carcass %in% c("B1", "B2", "B3", "B4", "B5"))
ch4cntrl <- subset_samples(ch4all, Carcass %in% c("C22", "C23", "C24", "C25", "C26"))

bleach_fam <- tax_glom(ch4bleach, taxrank = "Family")
control_fam <- tax_glom(ch4cntrl, taxrank = "Family")

```


venn diagram of overlapping families among sample types 
1400x800
```{r}

bleach_venn <- ggvenn_pq(bleach_fam, fact = "Order", taxonomic_rank = "Family", category.names = c("Soil Control", "Soil Underneath", "Coleoptera", "Carcass")) +
  scale_fill_distiller(palette = "BuPu", direction = 1)
bleach_venn

bleach_vennnew <- ggvenn_pq(bleach_fam, fact = "Order", taxonomic_rank = "Family", return_data_for_venn = TRUE)
```
barplot of carabidae and staph relative abundance over time
```{r}
relabdat <- data.frame(
  days = c(3,5,9), 
  carabidaeb =c(1, 0.73, 0), 
  carabidaec = c(0.93, 0.5, 0.19), 
  staphb =c(0, 0.091, 0.64), 
  staphc = c(0, 0.14, 0.76)
)

dat.melt <- reshape2::melt(relabdat, id.vars="days")

realplot <- ggplot(dat.melt, aes(x=days, y=value, fill=variable))+
  geom_col(position="dodge")+
  theme_bw()+
  scale_fill_brewer("Beetle Family and Treatment", palette="Set2", labels = c("Carabidae, B", "Carabidae, C", "Staphylinidae, B", "Staphylinidae, C"))+
  scale_x_continuous(labels = NULL, breaks = NULL)+
  labs(x=NULL, y="Relative Abundance")+
  facet_wrap(~days, scales="free_x")
realplot

```

